---
title: "NYPD Shooting Incident Data Report"
author: "Geoffrey Richards"
date: "3/23/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYPD Shooting Incident Data

This is a report summarizing and characterizing NYPD Shooting Data.  

The source of the data is https://catalog.data.gov/dataset.Several.

Columns filtered from data including INCIDENT_KEY, JURISDICTION_CODE, STATISTICAL_MURDER_FLAG, VIC_AGE_GROUP, VIC_SEX,VIC_RACE, X_COORD_CD, Y_COORD_CD, Lon_Lat, PERP_RACE, PERP_AGE_GROUP, and PRECINCT.

```{r, echo = TRUE, warning = FALSE, message = FALSE}
#Import libraries
library(tidyverse)
library(readr)
library(dplyr)

#Obtain data from source.
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
nypdShootingData = read_csv(url_in)

#Down select data.
nypdShootingData <- subset(nypdShootingData, select = -c(JURISDICTION_CODE))
nypdShootingData <- subset(nypdShootingData, select = -c(STATISTICAL_MURDER_FLAG))
nypdShootingData <- subset(nypdShootingData, select = -c(VIC_AGE_GROUP,VIC_SEX,VIC_RACE))
nypdShootingData <- subset(nypdShootingData, select = -c(Y_COORD_CD,Lon_Lat,PERP_RACE))
nypdShootingData <- subset(nypdShootingData, select = -c(PERP_AGE_GROUP,PRECINCT))
nypdShootingData <- subset(nypdShootingData, select = -c(INCIDENT_KEY,X_COORD_CD))

#Create tables for processing.
t <- table(floor(as.numeric(nypdShootingData$OCCUR_TIME, units="hours")/3600))
t2 <- table(nypdShootingData$BORO)
temp <- nypdShootingData$OCCUR_DATE
t3 <- table(as.numeric(substr(temp,nchar(temp) - 3,nchar(temp))))
t4 <- t3

#Fix data types.
nypdShootingData$OCCUR_TIME <- as.numeric(nypdShootingData$OCCUR_TIME, units="hours")
```
\newpage
NYPD Shooting Data Summary:
```{r nypdShootingData, echo = TRUE, warning = FALSE, message = FALSE}
#Produce summary.
summary(nypdShootingData)

```
\newpage
## Data Visualization

Interesting results reveal that most shootings occurred in the late evening and early morning.

```{r t, echo = TRUE, warning = FALSE, message = FALSE}
#Create bar plot of shooting incident times.
barplot(t,main="Time of Shooting",xlab="Hour",ylab="Count")
```
\newpage
Additionally, we can observe that Brooklyn is the most dangerous boro for which we have data.

```{r t2, echo = TRUE, warning = FALSE, message = FALSE}
#Create bar plot of shooting incidents by boro.
par(mar=c(8, 3, 3, 1))
barplot(t2,main="Boro Shooting Count",ylab="Count",las=2)
```
\newpage
## Analysis 1

By plotting the shooting counts by year, we can see a downward trend in shooting incidents.  However, we could note that there was a spike in shooting incidents in 2020. We might attribute this to the Covid lockdown and related social unrest.

This conclusion may be biased due to my own personal beliefs surrounding the Covid lockdown and personal experiences during that time period.  Additionally, the linear model may only capture a local trend in data.  It is possible that if we were to obtain data from as far back as the 1980s, we may actually see a global upward trend in shooting related incidents.  

Finally, the model is informationally biased because it not complex enough to establish a trend associated with shooting incidents.  We might expect, as an example, shooting related incidents to be descending rapidly, but then trail off asymptotically towards zero. Such is a typical trend for sociological variables. This complexity may require a nonlinear or higher order model.

```{r t3, echo = TRUE, warning = FALSE, message = FALSE}

#Fit linear regression model
Year = as.numeric(names(t3))
Count = as.numeric(t3)
lmod <- lm(Count~Year,data.frame(Year,Count))

#Obtain predictions
pred <- predict(lmod, newdata = data.frame(Year))

#Plot data
data<-data.frame(Year,Count,pred)
plot(data$Year,data$Count,type="l",main="Yearly Incidents",xlab="Year",ylab="Incidents")
lines(data$Year,data$pred,lty=2,lwd=2,col="green")
```


